---
output: pdf_document
header-includes: 
  - \usepackage[utf8]{inputenc}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage[margin=1in]{geometry}
  - \usepackage{amsmath,amssymb}
  - \usepackage{multicol}
  - \usepackage{array}
  - \usepackage{tabulary}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newcommand{\class}{EC 370}
\newcommand{\term}{Xiang LI (GE)}
\newcommand{\examnum}{Participation 8 (Chapter 9)}

\noindent
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r @{\extracolsep{6pt}} l}
\textbf{\class} & \textbf{Name: } & \makebox[2in]{\hrulefill}\\
\textbf{\term} &&\\
\textbf{\examnum} & \textbf{UO ID: } & \makebox[2in]{\hrulefill}\\
\end{tabular*}
\rule[2ex]{\textwidth}{2pt}

**INSTRUCTIONS**: Answer the following questions. Write legibly, take a high quality scan of your responses, compile all pages into one pdf, and upload only one pdf document to Canvas by **11:59pm (PDT), Sunday, June 7**. 

\vspace{.3in}
\textbf{QUESTION 1: T-account example I} 

Jane opens a checking account with a \$100 bill at the First National Bank, and the bank puts her \$100 bill into vault. 

_Instruction: To complete this balance sheets, show the **changes** of values for each item. For example, if you think the value of an item decreased by $10, write down -$10. If you think the value of an item increases by $10, write down +$10. If you think the value of an item isn't changed, leave it blank_

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library(kableExtra)
options(knitr.kable.NA = '')

dt <- data.frame(matrix(ncol = 2, nrow = 3))

colnames(dt)<-c("Assets", "Liabilities")

dt[1,] <- c("Reserves", "Deposits")
dt[2,] <- c("Securities", "Borrowings")
dt[3,] <- c("Loans", "Capital")

kable(dt, "latex", booktabs = T) %>%
  column_spec(1:2, width = "15em", border_right = T, border_left = T) %>%
  row_spec(0, bold=TRUE, align = "c") %>%
  kable_styling(font_size = 9, position = "center") 
```


\vspace{.3in}
\textbf{QUESTION 2: T-account example II} 

Jane opens a checking account at the First National Bank with a \$100 check written on an account at the Second National Bank. First National Bank deposits the check in its account at the Fed, and the Fed transfers \$100 of reserves from the Second National Bank to the First National Bank. Show the changes in the balance sheet.

_Instruction: To complete this balance sheets, show the **changes** of values for each item. For example, if you think the value of an item decreased by $10, write down -$10. If you think the value of an item increases by $10, write down +$10. If you think the value of an item isn't changed, leave it blank_

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library(kableExtra)
options(knitr.kable.NA = '')

dt <- data.frame(matrix(ncol = 4, nrow = 3))

colnames(dt)<-c("Assets", "Liabilities", "Assets", "Liabilities")

dt[1,] <- c("Reserves", "Deposits", "Reserves", "Deposits")
dt[2,] <- c("Securities", "Borrowings", "Securities", "Borrowings")
dt[3,] <- c("Loans", "Capital", "Loans", "Capital")

kable(dt, "latex", booktabs = T) %>%
  column_spec(1:4, width = "12em", border_right = T, border_left = T) %>%
  row_spec(0, bold=TRUE, align = "c") %>%
  kable_styling(font_size = 9, position = "center") %>%
  add_header_above(c("First National Bank" = 2, "Second National Bank" = 2), bold=TRUE, align = "c")
```

\newpage

\vspace{.3in}
\textbf{QUESTION 3: T-account example III} 

Suppose 10% required reserve ratio. First National Bank receives a $100 of checkable deposits. Then this bank put to productive use all excess reserves by making loans. Show the changes in the balance sheet.

_Instruction: To complete this balance sheets, show the **changes** of values for each item. For example, if you think the value of an item decreased by $10, write down -$10. If you think the value of an item increases by $10, write down +$10. If you think the value of an item isn't changed, leave it blank_

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library(kableExtra)
options(knitr.kable.NA = '')

dt <- data.frame(matrix(ncol = 2, nrow = 3))

colnames(dt)<-c("Assets", "Liabilities")

dt[1,] <- c("Reserves", "Deposits")
dt[2,] <- c("Securities", "Borrowings")
dt[3,] <- c("Loans", "Capital")

kable(dt, "latex", booktabs = T) %>%
  column_spec(1:2, width = "15em", border_right = T, border_left = T) %>%
  row_spec(0, bold=TRUE, align = "c") %>%
  kable_styling(font_size = 9, position = "center") 
```


\vspace{.3in}
\textbf{QUESTION 4: Sufficient Excess Reserves Example} 

The bank is required to keep 10% of deposits as reserves. Initial balance sheet is as follows. (Unit: millions)

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library(kableExtra)
options(knitr.kable.NA = '')

dt <- data.frame(matrix(ncol = 2, nrow = 3))

colnames(dt)<-c("Assets", "Liabilities")

dt[1,] <- c("Reserves: $20", "Deposits: $100")
dt[2,] <- c("Securities: $80", "Borrowings: $0")
dt[3,] <- c("Loans: $10", "Capital: $10")

kable(dt, "latex", booktabs = T) %>%
  column_spec(1:2, width = "15em", border_right = T, border_left = T) %>%
  row_spec(0, bold=TRUE, align = "c") %>%
  kable_styling(font_size = 9, position = "center") 
```

(1) How much excess reserves does this bank hold?

\vspace{1in}

(2) Suppose a deposit outflow of $10 million occurs. Update the balance sheet.

_Instruction: To complete the following balance sheets, do **NOT** show **changes** of values for each item. Instead, you should write **levels** of updated values for each item. For example, an item starts with value $100. If you think its value decreases by $10, write down $90 **instead of** -$10; if you think its value increases by $10, write down $110, **instead of** +$10; if you think its value isn't changed, write down $100_

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
library(kableExtra)
options(knitr.kable.NA = '')

dt <- data.frame(matrix(ncol = 2, nrow = 3))

colnames(dt)<-c("Assets", "Liabilities")

dt[1,] <- c("Reserves", "Deposits")
dt[2,] <- c("Securities", "Borrowings")
dt[3,] <- c("Loans", "Capital")

kable(dt, "latex", booktabs = T) %>%
  column_spec(1:2, width = "15em", border_right = T, border_left = T) %>%
  row_spec(0, bold=TRUE, align = "c") %>%
  kable_styling(font_size = 9, position = "center") 
```


(3) After deposit outflow occurs, does this bank meet the 10% required reserve requirement? If so, how much excess reserves does this bank hold?