---
title: "Ec 370 <br> Money and Banking"
subtitle: "Chapter 5: The Behavior of Interest Rate"
author: "Xiang LI"
date: "April 22, 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      countIncrementalSlides: false
      # countdown: 120000
    css: [default, metropolis, metropolis-fonts]
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(fig.width=6, fig.height=5) 
library(knitr)
library(tidyverse)
# library(gdtools)
# library(hrbrthemes)
library(plotly)
library(lubridate)
library(kableExtra)
library(htmlTable)
library(ggplot2)
library(grid)
```


<style>
  .large { font-size: 200% } 
  .small { font-size: 70% }
</style>

<style type="text/css">
.remark-slide-content {
    font-size: 25px;
    padding: 1em 1em 1em 1em;
}
</style>





# Today's Contents

- [Demand and Supply in the Bond Market](#Demand and Supply in the Bond Market)
- [Changes in Equilibrium Interest Rates](#Changes in Equilibrium Interest Rates)


---
class: inverse, center, middle
name: Motivation

# Demand and Supply in the Bond Market


<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>



---
# Demand Curve for Bond

- demand curve: relationship between quantity demanded and price
- x-axis: quantity demanded of bond
- y-axis: price of bond (**P**), also interest rate of bond (**i**)

--

Let’s consider purchasing a discount bonds
- no coupon payments
- face value (F): $1,000
- years to maturity: 1 year 
- holding period: 1 year


---
# Demand Curve for Bond

- What is interest rate (yield to maturity), if bond price is $950?

--
  - $P=\frac{F}{1+i}$
  - $i=\frac{F-P}{P}=\frac{1000-950}{950}=5.3\%$

--

- What is interest rate, if bond price is $850?

--
  - $i=\frac{F-P}{P}=\frac{1000-850}{850}=17.6\%$
  
--

- What is interest rate, if bond price is $750?

--
  - $i=\frac{F-P}{P}=\frac{1000-750}{750}=33.0\%$




---
# Demand Curve for Bond

- **negatively** related: if P $\uparrow$, then $B^d$ $\downarrow$

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
# slate <- "#314f4f"
# orange <- "#FFA500"

# line1 <- function(q) (-0.5*q+1000)
# x_range <- 100:500

# ggplot(
#   # data = data.frame(x=c(100,500)),
#   # aes(x=x)
# ) +
#   stat_function(aes(x_range), fun = line1, color = "#FFA500", size = 1)+
  #geom_segment(aes(x=300, y=850, xend=300, yend=750), color="#314f4f", linetype="dashed")+
  #geom_point(aes(300, 850), color = "#314f4f", size=4.5)+
  #geom_segment(aes(x=300, y=850, xend=100, yend=850), color="#314f4f", linetype="dashed")+
  # annotate(
  # geom = "text",
  # x = 300, y = 860,
  # label = "C",
  # size = 5,
  # hjust = 0,
  # vjust = 0,
  # color = "#314f4f")+
  #geom_segment(aes(x=100, y=950, xend=100, yend=750), color="#314f4f", linetype="dashed")+
  #geom_point(aes(100, 950), color = "#314f4f", size=4.5)+
  #geom_segment(aes(x=100, y=950, xend=90, yend=950), color="#314f4f", linetype="dashed")+
  # annotate(
  # geom = "text",
  # x = 100, y = 960,
  # label = "A",
  # size = 5,
  # hjust = 0,
  # vjust = 0,
  # color = "#314f4f")+
  #theme(axis.text.y = element_blank())+
  #annotate("text", label = "plot mpg vs. wt", x = -10, y = 930, size = 3, colour = "red")+
  # coord_cartesian(ylim = c(750, 950))+
  # coord_cartesian(xlim = c(100, 500))+
  # theme_classic()+
  # xlab("Quantity of Bonds, B ($ billions)")+
  # ylab("Price of Bonds, P ($)")


df1 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(100,300,500),
                     y=c(950,850,750))
    
ggplot(df1,aes(x=x,y=y)) +
  geom_line(color="blue")+
  annotate(geom = "text", x= 110, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 310, y=860, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 450, y=760, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 950), color = "blue", size=3)+
  geom_point(aes(300, 850), color = "blue", size=3)+
  geom_point(aes(500, 750), color = "blue", size=3)+
  annotate(geom = "text", x= 90, y=960, label = "A", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 490, y=760, label = "E", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 800, label = "B^d",parse = TRUE,size=5) #+
  # geom_segment(aes(x=100, y=950, xend=100, yend=750), color="#314f4f", linetype="dashed")+
  # geom_segment(aes(x=500, y=750, xend=100, yend=750), linetype="dashed")+
  # geom_segment(aes(x=300, y=850, xend=300, yend=750), linetype="dashed")+
  # geom_segment(aes(x=300, y=850, xend=100, yend=850), linetype="dashed")
```


---
# Supply Curve for Bond

- At lower P (higher i), quantity supplied of bonds is lower

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df2 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(500,300,100),
                     y=c(950,850,750))
    
ggplot(df2,aes(x=x,y=y)) +
  geom_line(color="blue")+
  annotate(geom = "text", x= 480, y=980, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 300, y=880, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 100, y=780, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 750), color = "blue", size=3)+
  geom_point(aes(300, 850), color = "blue", size=3)+
  geom_point(aes(500, 950), color = "blue", size=3)+
  annotate(geom = "text", x= 490, y=960, label = "I", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 90, y=760, label = "F", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 900, label = "B^s",parse = TRUE,size=5)
```





---
# Equilibrium 

- $B^d=B^s$: **bond market equilibirum**

--
  - equilibrium quantity: $300 billion
  - equilibrium price (P*): $850
  - equilibrium interest rate (i*): 17.6%
  


```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(100,300,500),
                     y=c(950,850,750))
df2 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(500,300,100),
                     y=c(950,850,750))

ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  annotate(geom = "text", x= 110, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 450, y=760, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 950), color = "blue", size=3)+
  geom_point(aes(500, 750), color = "blue", size=3)+
  annotate(geom = "text", x= 90, y=960, label = "A", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 490, y=760, label = "E", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 800, label = "B^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  annotate(geom = "text", x= 450, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 310, y=860, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color ="red") +
  annotate(geom = "text", x= 100, y=780, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 750), color = "blue", size=3)+
  geom_point(aes(300, 850), color = "red", size=3)+
  geom_point(aes(500, 950), color = "blue", size=3)+
  annotate(geom = "text", x= 490, y=960, label = "I", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate(geom = "text", x= 90, y=760, label = "F", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 900, label = "B^s",parse = TRUE,size=5)+
  geom_hline(yintercept=850, linetype="dashed", color = "red")
```
  


---
# Equilibrium 

- $B^d>B^s$: **excess demand**

--
  - P will rise until reaching P\*, i will fall until reaching i\*
  - move along supply curve from F to C, move along demand curve from E to C 
  
```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(100,300,500),
                     y=c(950,850,750))
df2 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(500,300,100),
                     y=c(950,850,750))

ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  annotate(geom = "text", x= 110, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 450, y=760, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "red") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 950), color = "blue", size=3)+
  geom_point(aes(500, 750), color = "red", size=3)+
  annotate(geom = "text", x= 90, y=960, label = "A", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 490, y=760, label = "E", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate('text', x = 450, y = 800, label = "B^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  annotate(geom = "text", x= 450, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3,color = "blue") +
  annotate(geom = "text", x= 310, y=860, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color ="blue") +
  annotate(geom = "text", x= 100, y=780, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "red") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 750), color = "red", size=3)+
  geom_point(aes(300, 850), color = "blue", size=3)+
  geom_point(aes(500, 950), color = "blue", size=3)+
  annotate(geom = "text", x= 490, y=960, label = "I", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 90, y=760, label = "F", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate('text', x = 450, y = 900, label = "B^s",parse = TRUE,size=5)+
  geom_hline(yintercept=750, linetype="dashed", color = "red")
```



---
# Equilibrium 

- $B^d<B^s$: **excess supply**

--
  - P will fall until reaching P\*, i will rise until reaching i\*
  - move along supply curve from I to C; move along demand curve from A to C 
  
```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(100,300,500),
                     y=c(950,850,750))
df2 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(500,300,100),
                     y=c(950,850,750))

ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  annotate(geom = "text", x= 110, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "red") +
  annotate(geom = "text", x= 450, y=760, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 950), color = "red", size=3)+
  geom_point(aes(500, 750), color = "blue", size=3)+
  annotate(geom = "text", x= 90, y=960, label = "A", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate(geom = "text", x= 490, y=760, label = "E", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 800, label = "B^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  annotate(geom = "text", x= 450, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "red") +
  annotate(geom = "text", x= 310, y=860, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color ="blue") +
  annotate(geom = "text", x= 100, y=780, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 750), color = "blue", size=3)+
  geom_point(aes(300, 850), color = "blue", size=3)+
  geom_point(aes(500, 950), color = "red", size=3)+
  annotate(geom = "text", x= 490, y=960, label = "I", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 90, y=760, label = "F", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 900, label = "B^s",parse = TRUE,size=5)+
  geom_hline(yintercept=950, linetype="dashed", color = "red")
```



---
# Equilibrium 

- The relationship between quantity demanded and quantity supplied of bonds determines interest rate 

--
- **move along** the demand/supply curve until reaching the **equilibirum interest rate**


```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(100,300,500),
                     y=c(950,850,750))
df2 <- data.frame(n=c("(i=5.3%)","(i=17.6%)","(i=33%)"),
                     x=c(500,300,100),
                     y=c(950,850,750))

ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  annotate(geom = "text", x= 110, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 450, y=760, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 950), color = "blue", size=3)+
  geom_point(aes(500, 750), color = "blue", size=3)+
  annotate(geom = "text", x= 90, y=960, label = "A", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 490, y=760, label = "E", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 800, label = "B^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  annotate(geom = "text", x= 450, y=960, label = "(i=5.3%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  annotate(geom = "text", x= 310, y=860, label = "(i=17.6%)", hjust = 0, vjust=0,size = 3, color ="red") +
  annotate(geom = "text", x= 100, y=780, label = "(i=33%)", hjust = 0, vjust=0,size = 3, color = "blue") +
  coord_cartesian(xlim = c(100, 500), clip = 'off') +
  theme(plot.margin = unit(c(1,3,1,1), "lines"))+
  theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  geom_point(aes(100, 750), color = "blue", size=3)+
  geom_point(aes(300, 850), color = "red", size=3)+
  geom_point(aes(500, 950), color = "blue", size=3)+
  annotate(geom = "text", x= 490, y=960, label = "I", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate(geom = "text", x= 290, y=860, label = "C", hjust = 0, vjust=0,size = 4, color = "red") +
  annotate(geom = "text", x= 90, y=760, label = "F", hjust = 0, vjust=0,size = 4, color = "blue") +
  annotate('text', x = 450, y = 900, label = "B^s",parse = TRUE,size=5)+
  geom_hline(yintercept=950, linetype="dashed")+
  geom_hline(yintercept=850, linetype="dashed", color="red")+
  geom_hline(yintercept=750, linetype="dashed")+
  geom_segment(aes(x =100, y = 930, xend = 100, yend = 870),
                  arrow = arrow(length = unit(0.5, "cm")), color="red")+
  geom_segment(aes(x =100, y = 770, xend = 100, yend = 830),
                  arrow = arrow(length = unit(0.5, "cm")), color="red")
```
 



<!---
* Recall Ec202: 
- **Ceteris paribus**: the assumption that all **other** economic variables are held **constant**
* So far, when we study how quantity demanded/supplied of bonds determines the change of interest rate
  + we have assumed all **other** economic variables that can affect interest rate are held **constant**
  + then, we **move along** demand/supply curve so that interest rate can change to its equilibrium level
- **Other** economic variables that can affect interest rate:
  + bond buyer's wealth
  
  
  + expected profitability of the corporate that issues the bond......
* What if these economic variables are **NOT constant**? 
* If one of those other economic variables that affect interest rate are **NOT** held constant, how will equilibrium interest rate change?
--->



---
class: inverse, center, middle
name: Motivation

# Changes in Equilibrium Interest Rates


<html><div style='float:left'></div><hr color='#EB811B' size=1px width=796px></html>




---
# Changes in Equilibrium Interest Rates

- Now, we will look at **shift** of supply/demand curves

--

- And how **shift** of supply/demand curves determines equilibrium bonds' prices and **equilibirum interest rate**



---
# Demand Curve Shift

Changes in factors that affect how many bonds people will want to hold will make **demand curve for bonds shift**
 
 
- Wealth
- Expected returns on bonds relative to alternative assets
  - Expected interest rate
  - Expected return on alternative assets
  - Expected inflation
- Risk of bonds relative to alternative assets
- Liquidity of bonds relative to alternative assets



---
# Demand Curve Shift
## (1) Wealth

- if wealth increases, demand for bonds at each bond price (or at each interest rate) rises
  - business cycle expansion
  - increased propensity to save



---
# Demand Curve Shift
## (1) Wealth

- **business cycle expansion (or: propensity to save increases)** $\Rightarrow$ wealth increases $\Rightarrow$ demand for bonds increases $\Rightarrow$ demand curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(x=c(1,5),y=c(5,1))
df2 <- data.frame(x=c(1,3),y=c(3,1))
  
demand_left <- ggplot() +
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  geom_line(aes(x=df2$x,y=df2$y), color="red")+
  coord_cartesian(xlim = c(1, 5), clip = 'off') +
  theme(axis.text.x=element_blank(), axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  #theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  annotate('text', x = 5, y = 1, label = "B[1]^d",parse = TRUE,size=5) +
  annotate('text', x = 3, y = 1, label = "B[2]^d",parse = TRUE,size=5)

demand_left
```


---
# Demand Curve Shift
## (1) Wealth

- if wealth decreases, demand for bonds at each bond price (or at each interest rate) decreases
  - business cycle recession
  - decreased propensity to save
  


---
# Demand Curve Shift
## (1) Wealth

- **business cycle recession (or: propensity to save decreases)** $\Rightarrow$ wealth decreases $\Rightarrow$ demand for bonds decreases $\Rightarrow$ demand curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_left
```




---
# Demand Curve Shift
## (2) Expected returns 

- Recall Chapter 4: if interest rate when you sell the bond is higher than the interest rate when you buy the bond $\Rightarrow$ the price at which you sell the bond is lower than the price at which you buy the bond 

--

- interest-rate risk: interest rate rises $\Rightarrow$ bond price falls $\Rightarrow$ capital loss $\Rightarrow$ rate of return falls


--

- if you expect that interest rate will rise in the future, then you also expect that return on bonds will fall; hence, you demand less bonds today 



---
# Demand Curve Shift
## (2) Expected returns 

- **expected future interest rates** increases $\Rightarrow$ expected return on bonds decreases $\Rightarrow$ demand for bonds decreases $\Rightarrow$ demand curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_left
```

- **expected future interest rates** decreases $\Rightarrow$ demand curve shifts to the right


---
# Demand Curve Shift
## (2) Expected returns 

- **expected return on alternative assets** increases $\Rightarrow$ expected returns on bonds relative to alternative assets decreases $\Rightarrow$ demand for bonds decreases $\Rightarrow$ demand curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis', fig.width=5, fig.height=4}
demand_left
```

- **expected return on alternative assets** decreases $\Rightarrow$ demand curve shifts to the right


---
# Demand Curve Shift
## (2) Expected returns 

- Recall Chapter 4: $r=i-\pi^e$

--

- **expected inflation** increases $\Rightarrow$ real interest rate decreases $\Rightarrow$ expected return on bonds decreases $\Rightarrow$ demand for bonds decreases $\Rightarrow$ demand curve shifts to the left



---
# Demand Curve Shift
## (2) Expected returns 

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_left
```

- **expected inflation** decreases $\Rightarrow$ demand curve shifts to right

---
# Demand Curve Shift
## (3) Risk 

- prices in the bond market become more volatile $\Rightarrow$ **riskiness of bonds** increases $\Rightarrow$ demand for bonds decreases $\Rightarrow$ demand curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_left
```



---
# Demand Curve Shift
## (3) Risk 

- an increase in the volatility of prices in another asset market (e.g. stock market)  $\Rightarrow$ **riskiness of alternative assets** increases $\Rightarrow$ demand for bonds increases $\Rightarrow$ demand curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(x=c(1,5),y=c(5,1))
df2 <- data.frame(x=c(1,3),y=c(3,1))
  
demand_right <- ggplot() +
  geom_line(aes(x=df1$x,y=df1$y), color="red")+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  coord_cartesian(xlim = c(1, 5), clip = 'off') +
  theme(axis.text.x=element_blank(), axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))+
  #theme_classic()+
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  annotate('text', x = 5, y = 1, label = "B[2]^d",parse = TRUE,size=5) +
  annotate('text', x = 3, y = 1, label = "B[1]^d",parse = TRUE,size=5)

demand_right
```

---
# Demand Curve Shift
## (4) Liquidity

- more people started trading in the bond market $\Rightarrow$ easier to sell bonds quickly $\Rightarrow$ **liquidity of bonds** increases $\Rightarrow$ demand for bonds increases $\Rightarrow$ demand curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_right
```



---
# Demand Curve Shift
## (4) Liquidity

- **liquidity of alternative assets** increases $\Rightarrow$ demand for bonds decreased $\Rightarrow$ demand curve shifts to the left
  - the reduction of brokerage commissions for trading common stocks in 1975

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
demand_left
```




---
# Supply Curve Shifts

Changes in factors that affect how many bonds corporates or government will want to issue will make **supply curve shift**
 
--
 
- Expected profitability of investment opportunities
- Expected inflation
- Government budget deficits




---
# Supply Curve Shifts
## (1) Expected Profitability of Investment Opportunities

- business cycle expansion $\Rightarrow$ **expected profitability of investment opportunities** increases $\Rightarrow$ supply of bonds increases $\Rightarrow$ supply curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(x=c(1,5),y=c(5,1))
df2 <- data.frame(x=c(1,5),y=c(1,5))
df3 <- data.frame(x=c(2,5),y=c(1,4))

supply_right <- ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  coord_cartesian(xlim = c(1, 5), clip = 'off') +
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  annotate('text', x = 5, y = 1, label = "B[1]^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="blue")+
  annotate('text', x = 5, y = 5, label = "B[1]^s",parse = TRUE,size=5)+
  geom_line(aes(x=df3$x,y=df3$y), color="red")+
  annotate('text', x = 5, y = 4, label = "B[2]^d",parse = TRUE,size=5)+
  theme(axis.text.x=element_blank(), axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

supply_right
```


---
# Supply Curve Shifts
## (1) Expected Profitability of Investment Opportunities

- business cycle recession $\Rightarrow$ **expected profitable investment opportunities** decreases $\Rightarrow$ supply of bonds decreases $\Rightarrow$ supply curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
df1 <- data.frame(x=c(1,5),y=c(5,1))
df2 <- data.frame(x=c(1,5),y=c(1,5))
df3 <- data.frame(x=c(2,5),y=c(1,4))

supply_left <- ggplot()+
  geom_line(aes(x=df1$x,y=df1$y), color="blue")+
  coord_cartesian(xlim = c(1, 5), clip = 'off') +
  xlab("Quantity of Bonds, B ($ billions)")+
  ylab("Price of Bonds, P ($)")+
  annotate('text', x = 5, y = 1, label = "B[1]^d",parse = TRUE,size=5)+
  geom_line(aes(x=df2$x,y=df2$y), color="red")+
  annotate('text', x = 5, y = 5, label = "B[2]^s",parse = TRUE,size=5)+
  geom_line(aes(x=df3$x,y=df3$y), color="blue")+
  annotate('text', x = 5, y = 4, label = "B[1]^d",parse = TRUE,size=5)+
  theme(axis.text.x=element_blank(), axis.text.y=element_blank(),axis.ticks.x=element_blank(),axis.ticks.y=element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"))

supply_left
```



---
# Supply Curve Shifts
## (2) Expected Inflation

- **expected inflation** increases $\Rightarrow$ real interest rate decreases $\Rightarrow$ real cost of borrowing decreases $\Rightarrow$ issuing more bonds $\Rightarrow$ supply of bonds increases $\Rightarrow$ supply curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
supply_right
```



---
# Supply Curve Shifts
## (2) Expected Inflation

- **expected inflation** decreases $\Rightarrow$ supply of bonds decreases $\Rightarrow$ supply curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
supply_left
```





---
# Supply Curve Shifts
## (3) Government Budget Deficits

- government’s expenditures is higher than its revenues $\Rightarrow$ **government deficits** $\Rightarrow$ U.S. Treasury issues bonds to finance government deficits $\Rightarrow$ supply of bonds increases $\Rightarrow$ supply curve shifts to the right

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
supply_right
```



---
# Supply Curve Shifts
## (3) Government Budget Deficits

- government’s expenditures is less than its revenues $\Rightarrow$ **government surpluses** $\Rightarrow$ supply of bonds decreases $\Rightarrow$ supply curve shifts to the left

--

```{r , echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
supply_right
```



<!---
library(webshot)
file_name <- paste0("file://", normalizePath("Apr22_Ch5_lecture_slide.html"))
webshot(file_name, "Apr22_Ch5_lecture_slide.pdf")
--->

