---
output: pdf_document
header-includes: 
  - \usepackage[utf8]{inputenc}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage[margin=1in]{geometry}
  - \usepackage{amsmath,amssymb}
  - \usepackage{multicol}
  - \usepackage{array}
  - \usepackage{tabulary}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newcommand{\class}{EC 370}
\newcommand{\term}{Xiang LI (GE)}
\newcommand{\examnum}{Participation 6 (Chapter 6)}

\noindent
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r @{\extracolsep{6pt}} l}
\textbf{\class} & \textbf{Name: } & \makebox[2in]{\hrulefill}\\
\textbf{\term} &&\\
\textbf{\examnum} & \textbf{UO ID: } & \makebox[2in]{\hrulefill}\\
\end{tabular*}
\rule[2ex]{\textwidth}{2pt}

**INSTRUCTIONS**: Answer the following questions. Write legibly, take a high quality scan of your responses, compile all pages into one pdf, and upload only one pdf document to Canvas by **11:59pm (PDT), Sunday, June 7**. 

\vspace{.3in}

\textbf{QUESTION 1} 

Higher default risk of corporate bonds relative to U.S. Treasury bonds increases the spread between the interest rates on these two bonds

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
library(ggplot2)
library(gridExtra)
library(grid)

blank1 <- ggplot()+
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4), clip = 'off') +
  theme_void()+
  geom_hline(yintercept=0)+
  geom_vline(xintercept = 0)+
  labs(caption="Corporate bonds") + 
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.2)))

blank2 <- ggplot()+
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4), clip = 'off') +
  theme_void()+
  geom_hline(yintercept=0)+
  geom_vline(xintercept = 0)+
  labs(caption="Treasury bonds") + 
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.2)))

grid.arrange(blank1, nullGrob(), blank2, widths=c(4,1,4))
```
	

\vspace{.5in}
	
\textbf{QUESTION 2} 

Lower liquidity of corporate bonds relative to Treasury bonds increases the spread between the interest rates on these two bonds 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
grid.arrange(blank1, nullGrob(), blank2, widths=c(4,1,4))
```

	
	
\newpage

\textbf{QUESTION 3} 

Favorable tax treatment of municipal bonds relative to Treasury bonds increases the spread between the interest rates on these two bonds 


```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
library(ggplot2)
library(gridExtra)
library(grid)

blank1 <- ggplot()+
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4), clip = 'off') +
  theme_void()+
  geom_hline(yintercept=0)+
  geom_vline(xintercept = 0)+
  labs(caption="Municipal bonds") + 
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.2)))

blank2 <- ggplot()+
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4), clip = 'off') +
  theme_void()+
  geom_hline(yintercept=0)+
  geom_vline(xintercept = 0)+
  labs(caption="Treasury bonds") + 
  theme(plot.caption = element_text(hjust=0.5, size=rel(1.2)))

grid.arrange(blank1, nullGrob(), blank2, widths=c(4,1,4))
```



	

