---
output: pdf_document
header-includes: 
  - \usepackage[utf8]{inputenc}
  - \usepackage{graphicx}
  - \usepackage{caption}
  - \usepackage{amsmath,amssymb}
  - \usepackage{multicol}
  - \usepackage{array}
  - \usepackage{tabulary}
geometry: "left=1in, right=1in, top=1in, bottom=1in"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newcommand{\class}{EC 370}
\newcommand{\term}{Xiang LI (GE)}
\newcommand{\examnum}{Participation 10 (Chapter 15)}

\noindent
\begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r @{\extracolsep{6pt}} l}
\textbf{\class} & \textbf{Name: } & \makebox[2in]{\hrulefill}\\
\textbf{\term} &&\\
\textbf{\examnum} & \textbf{UO ID: } & \makebox[2in]{\hrulefill}\\
\end{tabular*}
\rule[2ex]{\textwidth}{2pt}

**INSTRUCTIONS**: Answer the following questions. Write legibly, take a high quality scan of your responses, compile all pages into one pdf, and upload only one pdf document to Canvas by **11:59pm (PDT), Sunday, June 7**. 

\vspace{.3in}

\textbf{Graph 1: the impact on fed funds rate of an open market sale} 

(1) Left-hand-plot: vertical section of supply curve initially intersects demand curve at downward-sloped section

(2) Right-hand-plot: vertical section of supply curve initially intersects demand curve at flat section

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
library(ggplot2)
library(gridExtra)
library(grid)

blank <- ggplot()+
  coord_cartesian(xlim = c(0, 4), ylim = c(0, 4), clip = 'off') +
  theme_void()+
  geom_hline(yintercept=0)+
  geom_vline(xintercept = 0)

grid.arrange(blank, nullGrob(), blank, widths=c(4,1,4))
```


\textbf{Graph 2: the impact on fed funds rate of an increase in the discount rate}

(1) Left-hand-plot: the vertical section of the supply curve intersects the demand curve at the downward-sloped section

(2) Right-hand-plot: the flat section of the supply curve intersects the demand curve at the downward-sloped section

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
grid.arrange(blank, nullGrob(), blank, widths=c(4,1,4))
```

	
	
\newpage

\textbf{Graph 3: the impact on fed funds rate of a decrease in the required reserve ratio}

The vertical section of the supply curve intersects the demand curve at the downward-sloped section

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", out.width="50%", fig.width=5, fig.height=4}
blank
```


\textbf{Graph 4: the impact on fed funds rate of a decrease in the interest rate on reserves}

(1) Left-hand-plot: the vertical section of the supply curve intersects the demand curve at the downward-sloped section

(2) Right-hand-plot: the vertical section of the supply curve intersects the demand curve at the flat section

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.show="hold", fig.width=7, fig.height=2.5, fig.align='center'}
grid.arrange(blank, nullGrob(), blank, widths=c(4,1,4))
```
	

	
	
	





